---
title: "my title"
author: "Dummy Surname (dummy@mail.com)"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`"      
output:
  html_document:
    toc: true
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Analysis of the Heart Disease Dataset 
Load the data from
[here](https://raw.githubusercontent.com/jpinero/DMI_2021/main/datasets/heart_disease_dataset.csv), and the description is [here](https://raw.githubusercontent.com/jpinero/DMI_2021/main/datasets/heart_disease_description.txt). 
The original dataset comes from [here](https://archive.ics.uci.edu/ml/datasets/Heart+Disease) and corresponds to the [processed cleveland data](https://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/processed.cleveland.data)

## Perform an EDA on the dataset

```{r}
library(ggplot2)

# Read table
data <- read.table("data/heart_disease_dataset.csv", header=TRUE, na.strings = c("?"))

# Summarize the table
summary(data)

# Check duplicates
data$patient_id[duplicated(data$patient_id)] # No duplicates
data$patient_id = NULL # Remove as it won't be of any help to our data

# Change numerical values to categories
# Sex
data$sex <- factor(ifelse(data$sex == 1, "male", "female"))
# Chest pain type
data$cp <- factor(data$cp, levels = c(1,2,3,4), labels = c("typical angina", "atypical angina","non-anginal pain","asymptomatic"))
# Fasting blood sugar
data$fbs <- factor(ifelse(data$fbs == 1, "TRUE", "FALSE"))
# Resting electrocardiographic results
data$restecg <- factor(data$restecg, levels = c(0,1,2), labels = c("normal", "ST-T wave abnormality","Ventricular hypertrophy"))
# Exercise induced angina
data$exang <- factor(ifelse(data$exang == 0, "No", "Yes"))
# Slope of the peak exercise ST segment
data$slope <- factor(data$slope, levels = c(1,2,3), labels = c("Upsloping", "Flat","Downslopping"))
# "thal" variable
data$thal <- factor(data$thal, levels = c(3,6,7), labels = c("Normal", "Fixed defect","Reversable defect"))
# Diagnosis of heart disease
data$num <- factor(ifelse(data$num > 0, "Disease", ifelse(data$num == 0, "No disease", data$num)))
# Transform "ca" (number of major vessels) into a categorical variable
data$ca <- as.factor(data$ca)
# Then, we check the variables
str(data)

# Handling missing data using mice
library(mice)
imputed_data <- mice(data)
# Store the completed data into our main table
data <- complete(imputed_data)

# Handling outliers
library(cowplot)
library(ggplotify)
# For each numeric variable, we will be showing a boxplot
age_bp <- as.ggplot(~boxplot(data$age))
trestbps_bp <- as.ggplot(~boxplot(data$trestbps))
chol_bp <- as.ggplot(~boxplot(data$chol))
thalach_bp <- as.ggplot(~boxplot(data$thalach))
oldpeak_bp <- as.ggplot(~boxplot(data$oldpeak))

grid.arrange(age_bp, trestbps_bp, chol_bp, thalach_bp, oldpeak_bp)

# We can see some outliers in all the variables except age
# Lets see the outliers
summary(data[c('age','trestbps','chol','thalach','oldpeak')])
# Using the IQR method, we can see that trestbps, chol, and oldpeak have outliers as they are above the bounds of Q1 and Q3

# Using the outliers package, we will check these variables by their z-score
library(outliers)
# Save to new data
new_data <- data
# Score them and get there index, if z-score is higher than 3 or lower than -3, then it is considered an outlier

# trestbps
z_scores <- scores(new_data$trestbps, type = "z")
outlier_indices <- which(z_scores > 3 | z_scores < -3)
# Accomodation: Make equal to the next highest value
max_non_outlier <- max(new_data$trestbps[-outlier_indices])
new_data$trestbps[outlier_indices] <- max_non_outlier

# Chol
z_scores <- scores(new_data$chol, type = "z")
outlier_indices <- which(z_scores > 3 | z_scores < -3)
# Accomodation: Make equal to the next highest value
max_non_outlier <- max(new_data$chol[-outlier_indices])
new_data$chol[outlier_indices] <- max_non_outlier

# oldpeak
z_scores <- scores(new_data$oldpeak, type = "z")
outlier_indices <- which(z_scores > 3 | z_scores < -3)
# Accomodation: Make equal to the next highest value
max_non_outlier <- max(new_data$oldpeak[-outlier_indices])
new_data$oldpeak[outlier_indices] <- max_non_outlier

library(DataExplorer)
# Using the DataExplorer package, we can now get a report of our new data for univariate EDA
DataExplorer::create_report(new_data)

```

## Create visualizations in order to show which variables seem to be more associated with heart disease

```{r}
library(dplyr)
library(tidyr)
# We will start by Bivariate anaylsis of the 'num' variable (Diagnosis of heart disease)

# 1. Categorical variables
new_data[, c("sex", "fbs", "exang", "ca", "cp", "restecg", "slope", "thal", "num")] %>% gather(cols, value, -num) %>% ggplot(aes(x = value, fill = num)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ cols, scales = "free", labeller = labeller(cols = c(
    sex = "Gender (Male/Female)",
    fbs = "Fasting Blood Sugar > 120 mg/dl",
    exang = "Exercise Induced Angina",
    ca = "Number of Major Vessels (0-3)",
    cp = "Chest pain type",
    restecg = "Resting electrocardiographic results",
    slope = "Slope of the peak exercise ST segment",
    thal = "Thalassemia"
  ))) +
  scale_fill_manual(values = c("red", "lightgreen"), labels = c("Disease", "No Disease")) +
  labs(fill = "Heart Disease") +
  xlab("Levels") + ylab("Frequency") +
  ggtitle("Comparison of categorical variables\n for diagnosis of heart disease") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))

# 2. Continous variables
new_data[, c("age", "chol", "oldpeak", "thalach", "trestbps", "num")] %>% 
  gather(cols, value, -num) %>% 
  ggplot(aes(x = num, y = value, fill = num)) +
  geom_violin(trim = FALSE, alpha = 0.7) + 
  facet_wrap(~ cols, scales = "free", labeller = labeller(cols = c(
    age = "Age",
    chol = "Serum Cholesterol (mg/dl)",
    oldpeak = "ST Depression (Exercise relative to rest)",
    thalach = "Max Heart Rate Achieved",
    trestbps = "Resting Blood Pressure (mm Hg)"
  ))) +
  scale_fill_manual(values = c("red", "lightgreen"), labels = c("Disease", "No Disease")) +
  labs(fill = "Heart Disease", x = "Heart Disease Status", y = "Values") +
  ggtitle("Comparison of Continuous Variables\n for Diagnosis of Heart Disease") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```


# 2 Difference in mortality rates in hospitalized COVID-19 patients 
Using the supplementary material from the [Difference in mortality rates in hospitalized COVID-19 patients identified by cytokine profile clustering using a machine learning approach: An outcome prediction alternative](https://www.frontiersin.org/articles/10.3389/fmed.2022.987182/full), perform the following tasks

## Reproduce Figure 1 from the publication

```{r}
library(ggplotify)
library(readxl)
Table_1 <- read_excel("~/Desktop/OneDrive/Term2/DMI/HandsOn/1/Article/Table 1.XLSX", skip = 1, na = "NI")

#summary(Table_1)
# Fixing errors in the data
Table_1 <- Table_1[!duplicated(Table_1$ID), ] # Remove duplicated ID records

Table_1$Gender <- gsub(72, NA, Table_1$Gender)

Table_1$BMI <- gsub(",", ".", Table_1$BMI) # Change comma to decimal
Table_1$BMI <- gsub(0, NA, Table_1$BMI) # Make 0 BMI a NA value
Table_1$BMI <- as.numeric(Table_1$BMI)

Table_1[Table_1$ID == 'COAG-HSJD-180', "Admission to ICU"] <- "Yes" # Incorrect addition of a random number before Yes
Table_1[Table_1$ID == 'COAG-HSJD-076', "Admission to ICU"] <- NA # Incorrect data

Table_1[!Table_1$`ARDS Diagnosis` %in% c("Yes", "No"), "ARDS Diagnosis"] <- NA # Incorrect data

Table_1$`Use of NIV` <- gsub(8, NA, Table_1$`Use of NIV`)

Table_1$Death <- gsub(3, NA, Table_1$Death)

library(patchwork)
library(ggplot2)
# Replicate figure A
plot_A <- as.ggplot(~hist(
  Table_1$Age, 
  ylim = c(0, 50), 
  main = "Age", 
  xlab = "Age (years)", 
  ylab = "Frequency (n)", 
  col = "lightcyan2" #Plot A
))

# Replicate figure B
library(gridExtra)
library(grid)
library(ggplotify)

grid_table <- data.frame(
  Clinical= c('G1', 'G2', 'G3', 'G4'),
  NIV = c('-', '-/+', '+', '-/+'),
  AMV = c('-', '+/-', '-', '+'),
  ARDS = c('-', '-', '+', '+')
)
colnames(grid_table)[1] = 'Clinical\nclassification'
theme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 1.0)),
    colhead = list(fg_params=list(cex = 0.8)),
    rowhead = list(fg_params=list(cex = 0.8)))
grob <- tableGrob(grid_table, theme = theme, rows = NULL)
grob$widths <- unit(rep(0.6 / ncol(grob), ncol(grob)), "npc")
grob$heights <- unit(rep(0.6 / nrow(grob), nrow(grob)), "npc") # new
plot_B <- as.ggplot(grob) + ggtitle("Definition of the clinical classification") + theme(plot.title = element_text(hjust = 0.5, face = "bold")) 

# Replicate figure C
assign_group <- function(NIV, AMV, ARDS) {
  if (is.na(NIV) || is.na(AMV) || is.na(ARDS)) {
    return(NA)
  }
  if (NIV == "No" && AMV == "No" && ARDS == "No") {
    return("G1")
  } else if ((NIV %in% c("No", "Yes")) && (AMV %in% c("No", "Yes")) && ARDS == "No") {
    return("G2")
  } else if (NIV == "Yes" && AMV == "No" && ARDS == "Yes") {
    return("G3")
  } else if ((NIV %in% c("No", "Yes")) && AMV == "Yes" && ARDS == "Yes") {
    return("G4")
  } else {
    return(NA)
  }
}

Table_1$Group <- mapply(assign_group, Table_1$`Use of NIV`, Table_1$`Use of AMV`, Table_1$`ARDS Diagnosis`)

bar_tb <- table(Table_1$Group)
bar_df <- data.frame(
  names = names(bar_tb),
  numbers = as.numeric(bar_tb)
)
plot_C <- function() {
  plot_C <- barplot(
    names = bar_df$names,
    height = bar_df$numbers,
    ylim = c(0, 80),
    main = "Clinical classification",
    xlab = "Clinical classification",
    ylab = "Frequency (n)",
    col = c("aquamarine2", "khaki1", "plum2", "indianred1"))
  
  text(plot_C, bar_df$numbers + 4, labels = bar_df$numbers, cex = 1)
}
plot_C <- as.ggplot(~plot_C())

bar_tb_d <- table(Table_1$Death)
bar_df_d <- data.frame(
  names = names(bar_tb_d),
  numbers = as.numeric(bar_tb_d)
)

plot_D <- function() {
  plot_D <- barplot(
    names = bar_df_d$names,
    height = bar_df_d$numbers,
    ylim = c(0, 150),
    xlab = "Death",
    ylab = "Frequency (n)",
    col = c("aquamarine2", "khaki1"),
    main = "Vital status",
    yaxt = "n"
  )
  text(plot_D, bar_df_d$numbers + 6, labels = bar_df_d$numbers, cex = 1)
  axis(2, at = seq(0, 150, by = 50), labels = seq(0, 150, by = 50))
}
plot_D <- as.ggplot(~plot_D())

wrap_elements(panel = plot_A, clip = FALSE) + 
  plot_B +
  wrap_elements(panel = plot_C, clip = FALSE) +
  wrap_elements(panel = plot_D, clip = FALSE) + 
  plot_annotation(tag_levels = 'A') # Tag plots
```

## Reproduce Figure 2 from the publication
but instead of representing the clusters in the annotation, represent the groups (G1 to G4)

```{r}

```

## Improve figure 2 of the publication
Add a second annotation with information of deathm and a third one with information of gender

```{r}

```


# session info {.unnumbered}

```{r, results='asis',  echo=FALSE, message=FALSE }
sessionInfo()
```
